---
title: "HATCO Dataset"
author: "Devrim Nesipoglu - BDA 503 - Fall 2017 - Assignment Two"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


## Multivariate Data Analysis

HATCO provides a simplified set of variables amenable to all of the basic multivariate techniques. It was obtained from the Hair, Anderson, and Tatham Company (HATCO), a large (though nonexistent) industrial supplier. It represents surveys of HATCO customers that were collected through an established marketing research firm.

The dataset consists of 100 observations on 14 separate variables and is an example of a segmentation study for a business-to-business situation, specifically a survey of existing customers of HATCO.


## Data Dictionary

Three types of information were collected:

* First is the perception of HATCO on seven attributes identified in past studies.The respondents, purchasing managers of firms buying from HATCO, rated HATCO on each attribute.

* Second are actual purchase outcomes, either the evaluations of each respondent's satisfaction with HATCO.

* Third are general characteristics of the purchasing companies.

Perceptions of HATCO

- X1 Delivery speed -- amount of time it takes to deliver the product once an order has been confirmed

- X2 Price level -- perceived level of price charged by product suppliers

- X3 Price flexibility -- perceived willingness of HATCI representatives to negotiate price

- X4 Manufacturer's image

- X5 Service

- X6 Salesforce's image

- X7 Product quality

Purchase Outcomes: 

- X9 Usage Level : measured on a 100 point percentage scale, ranging from 0 to 100 percent.

- X10 Satisfaction Level 

Purchaser Characteristics

- X8 Size of firm 

- X11 Specification buying

- X12 Structure of procurement

- X13 Type of industry

- X14 Type of buying situation :  This variable has three categories: 1 = new task, 2 = modified rebuy, 3 = straight rebuy
    
    

```{r include=TRUE}

#Read the data and omit the first row
library(readxl)
HATCO <- read_excel("C:/Users/devrim/Desktop/BDA503/Personal Dataset/HATCO.xlsx")
View(HATCO)
```

Examine number of rows, columns

```{r include= TRUE}

#Number of Rows
nrow(HATCO)

#Number of cols
nrow(HATCO)

```


## Data structure of data

```{r }
#structure of data
str(HATCO)

#summary of data
summary(HATCO$x1)

```

## Data analysis

"Sapply"" is used to analysze more than one vector in this dataset


```{r}
colMeans(HATCO[sapply(HATCO,is.numeric)])
```

Mean is used for taking average of only one variable.


```{r}
mean(HATCO$x2)
```


Values between second and fifth columns are calculated. Start value is 2.


```{r}
lapply(HATCO[,2:5],mean)
```

## Data Exploration and Visualization



- 


```{r}
#Analyzing X1 variable with box plot

boxplot(HATCO$x1)

#X8 Size of firm. It has two catefories, 1 is used for large and 0 is used for small firms.
#X1 is Delivery speed. According to data, lets look the effect of X1 on X8.

boxplot(HATCO$x1~HATCO$x8,col=rainbow(7))

#Here we can say that delivery speed is of large scale firms are better than small firms. Large firms deliver average time is 2.5 min. For red one is 4 min
```

Trends of sales in HATCO

```{r}
library(ggplot2)

# First lets analyze if variable is distributed normal with using ggplot

qqplot(HATCO$x1,HATCO$id)

# Hereby we can see that if we omit first datas, it will be distributed more lineer. Lets sort them.

```

Sorting data to ensure more linear visualization and normal distribution

```{r}
HATCO$y <- sort(HATCO$x1, decreasing=FALSE)
HATCO1 <- HATCO[order(HATCO$x1),]
HATCO2 <- HATCO1[20:100,]
qqplot(HATCO2$x1,HATCO$id)
#After sorting data, we have more lineer visualization and normal distribution.
```

Lets look frequency of data by using plyr and bar plot.

```{r}
library(plyr)
barplot(count(HATCO$x8)$freq)
count(HATCO$x8)$x
```


```{r}
y <- count(HATCO$x8)
class(y)
barplot(y$freq)
y$x

library(plyr)
barplot(count(HATCO$x14)$freq, col=rainbow(7))
barplot(y$freq,names.arg =y$x,main = "Frequency table of company")
```

